---
title: "Covid91 vaccine study Final2022"
author: "Kimberly Lemus Munoz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

```




# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males
```{r}
Males <- filter(FinalDat, sex=="M")
```

```{r}
Males <- select(Males, -sex)
```

## Graphic Descriptive Results
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


The bar chart shows that males who take the vaccine (covid91) are far less likely to contract the virus covid19. Males who take the placebo are 68% more likely to contract covid19 versus males who took the vaccine and contacted covid19 32% of the time.

## Numerical Summaries

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```  


The percentage of infection is much higher for males who took the placebo. Those who took the placebo became infected approximately 68%  of the time versus 32% who took the vaccine and contracted covid19. The Males in the placebo group are 3 times more likely to become infected when receiving the placebo.These results are indicating that the vaccine is effective.



## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fishtestm<-fisher.test(table1)
print(fishtestm)
```  

The inferential results show that those that took the placebo are a little over twice as likely to catch covid19 as those that took the vaccine. The p-value being very low signifies that there is a small chance of getting these results if there was no difference between the placebo and the vaccine.


## Efficecy
```{r}
options(digits = 1)
maleffic=fishtestm$estimate/(1+fishtestm$estimate)*100
malefficlo=fishtestm$conf.int[1]/(1+fishtestm$conf.int[1])*100
maleffichi=fishtestm$conf.int[2]/(1+fishtestm$conf.int[2])*100
print(maleffic)
print(malefficlo)
print(maleffichi)
```
The confidence interval for the effectiveness of the vaccine on males has a lower `r malefficlo` and a high  `r maleffichi`.

# Females
```{r}
Females <- filter(FinalDat, sex=="F")
```

```{r}
Females <- select(Females, -sex)
```
## Graphic Descriptive Results


```{r}

dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```  

The bar chart shows that Females who take the vaccine (covid91) are less likely to contract the virus covid19. Females who took the placebo are 58% more likely to contract covid19 versus Females who took the vaccine and contacted covid19 40% of the time.


## Numerical Summaries
```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```  
The percentage of infection is much higher for Females who took the placebo. Those who took the placebo became infected approximately 59%  of the time versus 41% who took the vaccine and contracted covid19. 


## Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fishtestf<-fisher.test(table2)
fishtestf
```  

 The results show about a 1 odds ratio. This tells us that those that took the placebo are about 1 times as likely to catch covid19 than those that took the vaccine. The results also gave a low p-value which indicates a low chance of getting these results if there was no difference between females who took the placebo or the vaccine.

## Efficecy
```{r}
femeffic=fishtestf$estimate/(1+fishtestf$estimate)*100
femefficlo=fishtestf$conf.int[1]/(1+fishtestf$conf.int[1])*100
femeffichi=fishtestf$conf.int[2]/(1+fishtestf$conf.int[2])*100
print(femeffic)
print(femefficlo)
print(femeffichi)
```
The confidence interval of the effectiveness of the vaccine for females has a low of `r femefficlo` and a high of `r femeffichi`.


# LGBTQ

```{r}
LGBTQ = filter(FinalData,LGBTQ=="gay")

```




## Graphic Descriptive Results

```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")

```  


The bar chart shows that in the individuals in subgroup LGBTQ  who take the vaccine (covid91) are more likely to contract the virus covid19. Subgroup LGBTQ who take the placebo are 13% less likely to contract covid19 versus LGBTQ subgroup who took the vaccine and contacted covid19 87% of the time. 


## Numerical Summaries

```{r}
table3<- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```  


The percentage of infection is higher for subgroup LGBTQ who took the vaccine. Those who took the placebo became infected approximately 13%  of the time versus 87% who took the vaccine and contracted covid19. The subgroup LBGTQ in the vaccine group are 3 times more likely to become infected when receiving the vaccine.


## Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fishtestL<-fisher.test(table3)
fishtestL
```    

The odds ratio of 0.1 tells us that those that took the placebo are 0.1 times as likely to catch covid19 than those that took the vaccine.

## Efficecy

```{r}
Leffic=fishtestL$estimate/(1+fishtestL$estimate)*100
Lefficlo=fishtestL$conf.int[1]/(1+fishtestL$conf.int[1])*100
Leffichi=fishtestL$conf.int[2]/(1+fishtestL$conf.int[2])*100
print(Leffic)
print(Lefficlo)
print(Leffichi)
```
The efficacy shows that the vaccine is 13% effective. The confidence interval for the effectiveness of the vaccine shows a low of `r Lefficlo` and a high of `r Leffichi`.



# Druggies

```{r}
DrugUser<-filter(FinalData, DrugUser == "yes")
```

## Graphic Description
```{r}
dd2 <- DrugUser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The results from the barchart for the Drug users shows that those that took the vaccine got covid19 more than those that took the placebo. About 85% of those that got covid19 took the vaccine, while about 15% of those that got it took the placebo

## Numerical Results
```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

The numerical results tell us that for the drug users about 3% of those that took the vaccine got covid19, and 0.6% of those who took the placebo got covid19.


## Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fishtestD<-fisher.test(table4)
fishtestD
```

The inferential results showed that those that took the placebo were 0.2 times as likely to get covid than those that got the vaccine. The p-value is very low which tells us that there is very little chance of getting these results if there was no difference between the placebo and the vaccine for drug users.


## Efficacy
```{r}
Deffic=fishtestD$estimate/(1+fishtestD$estimate)*100
Defficlo=fishtestD$conf.int[1]/(1+fishtestD$conf.int[1])*100
Deffichi=fishtestD$conf.int[2]/(1+fishtestD$conf.int[2])*100
print(Deffic)
print(Defficlo)
print(Deffichi)
```

The efficacy shows that the vaccine is 15% effective. This confidence interval has a low of `r Defficlo` and a high of `r Deffichi`.








# Discussion and Conclusion

The immunization for the males was successful in shielding them from covid19. The viability of the vaccine for the males was `r maleffic`, which was the most elevated for any gathering. The vaccine was good in keeping females from getting covid19 too. The viability results showed that for females the vaccine has a `r femeffic` adequacy. The vaccine didn't safeguard the people that were LGBTQ from getting covid19. The adequacy of the vaccine on the LGBTQ people was just `r Leffic`, which was the most minimal of any gathering. The immunization didn't work for the DrugUsers either, with just `r Deffic` viability.

The vaccine ought  be taken on the off chance that a person is male or female with no other variables. It isn't recommended to take the vaccine on the off chance that you are a person who is LGBTQ or a Drug User.



