---
title: "Covid91 vaccine study Final2022"
author: "Kimberly Lemus Munoz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

```

```{r}
FinalDat <- read.csv(file="final_Data.csv")
```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
Males <- filter(FinalDat, sex=="M")
```

```{r}
Males <- select(Males, -sex)
```

## Graphic Descriptive Results
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
The bar chart shows that males who take the vaccine (covid91) are far less likely to contract the virus covid19. Males who take the placebo are 68% more likely to contract covid19 versus males who took the vaccine and contacted covid19 32% of the time. 
## Numerical Summaries

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```  
The percentage of infection is much higher for males who took the placebo. Those who took the placebo became infected approximately 67.78%  of the time versus 32.22% who took the vaccine and contracted covid19. The Males in the placebo group are 2 times more likely to become infected when receiving the placebo.



## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  

I reject the  Null Hypothesis, and accept the Alternative hypothesis as true. There is a significant negative effect when Males took the placebo vs the covid91 vaccine which has a positive effect.  The Chi-squared test shows that the P-value is 0.00000000004812 an amount significantly less than .05. The Fisher exact test shows the P-Value is 0.00000000002645 also an number significalty less than 0.05. The test also confirms that the odds ratio of men who take the placebo contract covid19 2 to 1 more times. The covid91 vaccine is effective for the Male subgroup tested. I would recommend the Covid91 vaccine for the Male subgroup. 





## Females
## Graphic Descriptive Results
## Numerical Summaries
## Inferential Results

## LGBTQ
## Graphic Descriptive Results
## Numerical Summaries
## Inferential Results


## Druggies
## Graphic Descriptive Results
## Numerical Summaries
## Inferential Results


#Discussion and Conclusion

